---
title: "Discussion Week 4"
author: "Julia Ferris"
date: "2024-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question M20 Page 403

### The set B = {v1, v2, v3, v4} is a basis of the vector space P3, polynomials with degree 3 or less. Therefore ρB is a linear transformation, according to Theorem VRLT. Find a "formula" for ρB. In other words, find an expression for ρB(a + bx + cx\^2 + dx\^3).

#### v1 = 1 - 5x - 22x\^2 + 3x\^3

#### v2 = -2 + 11x + 49x\^2 - 8x\^3

#### v3 = -1 + 7x + 33x\^2 - 8x\^3

#### v4 = -1 + 4x + 16x\^2 + x\^3

### Step 1

Determine the meaning of the question. In this question, ρB is a linear transformation going from P3 to $C^4$. The basis of $C^4$ can be defined as {1, x, $x^2$, $x^3$}.

### Step 2

Determine the vector representations of the elements of $C^4$. This is done by solving the systems of equations from the set B.

I will show the set up for solving each system of equations and use R to come to each of the results.

C = {1, x, $x^2$, $x^3$}, which can also be represented in vectors as 1{1,0,0,0}, x{0,1,0,0,}, $x^2${0,0,1,0}, and $x^3${0,0,0,1}.

The matrix used on the left side of the following equations is based on the original vectors from the question.

For 1{1,0,0,0}:

$\begin{bmatrix} 1 & -2 & -1 & -1 \\ -5 & 11 & 7 & 4 \\ -22 & 49 & 33 & 16 \\ 3 & -8 & -8 & 1 \end{bmatrix}$ = $\begin{bmatrix} 1 \\ 0 \\ 0 \\ 0 \end{bmatrix}$

Row reduce to solve the system of equations:

$\begin{bmatrix} 1 & -2 & -1 & -1 & 1\\ -5 & 11 & 7 & 4 & 0\\ -22 & 49 & 33 & 16 & 0 \\ 3 & -8 & -8 & 1 & 0 \end{bmatrix}$

```{r}
library(pracma)
A <- matrix(c(1, -2, -1, -1, 1, -5, 11, 7, 4, 0, -22, 49, 33, 16, 0, 3, -8, -8, 1, 0), 4, 5, byrow = TRUE)
part1 <- rref(A)[,5]      
```

For x{0,1,0,0}:

$\begin{bmatrix} 1 & -2 & -1 & -1 \\ -5 & 11 & 7 & 4 \\ -22 & 49 & 33 & 16 \\ 3 & -8 & -8 & 1 \end{bmatrix}$ = $\begin{bmatrix} 0 \\ 1 \\ 0 \\ 0 \end{bmatrix}$

$\begin{bmatrix} 1 & -2 & -1 & -1 & 0\\ -5 & 11 & 7 & 4 & 1 \\ -22 & 49 & 33 & 16 & 0 \\ 3 & -8 & -8 & 1 & 0 \end{bmatrix}$

```{r}
A <- matrix(c(1, -2, -1, -1, 0, -5, 11, 7, 4, 1, -22, 49, 33, 16, 0, 3, -8, -8, 1, 0), 4, 5, byrow = TRUE)
part2 <- rref(A)[,5]    
```

For x{0,0,1,0}:

$\begin{bmatrix} 1 & -2 & -1 & -1 \\ -5 & 11 & 7 & 4 \\ -22 & 49 & 33 & 16 \\ 3 & -8 & -8 & 1 \end{bmatrix}$ = $\begin{bmatrix} 0 \\ 0 \\ 1 \\ 0 \end{bmatrix}$

$\begin{bmatrix} 1 & -2 & -1 & -1 & 0\\ -5 & 11 & 7 & 4 & 1 \\ -22 & 49 & 33 & 16 & 0 \\ 3 & -8 & -8 & 1 & 0 \end{bmatrix}$

```{r}
A <- matrix(c(1, -2, -1, -1, 0, -5, 11, 7, 4, 0, -22, 49, 33, 16, 1, 3, -8, -8, 1, 0), 4, 5, byrow = TRUE)
part3 <- rref(A)[,5]    
```

For x{0,0,0,1}:

$\begin{bmatrix} 1 & -2 & -1 & -1 \\ -5 & 11 & 7 & 4 \\ -22 & 49 & 33 & 16 \\ 3 & -8 & -8 & 1 \end{bmatrix}$ = $\begin{bmatrix} 0 \\ 0 \\ 0 \\ 1 \end{bmatrix}$

$\begin{bmatrix} 1 & -2 & -1 & -1 & 0 \\ -5 & 11 & 7 & 4 & 0 \\ -22 & 49 & 33 & 16 & 1 \\ 3 & -8 & -8 & 1 & 0 \end{bmatrix}$

```{r}
A <- matrix(c(1, -2, -1, -1, 0, -5, 11, 7, 4, 0, -22, 49, 33, 16, 0, 3, -8, -8, 1, 1), 4, 5, byrow = TRUE)
part4 <- rref(A)[,5] 
```

These are the vectors found during this step in a matrix where each column is one vector.

```{r}
matrix(c(part1, part2, part3, part4), 4, 4, byrow = FALSE)
```

### Step 3

Create the formula to solve the problem. Our final expression will be in the form ρB(a + bx + $cx^2$ + $dx^3$). 

This formula is equivalent to aρB(1) + bρB(x) + cρB($x^2$) + dρB($x^3$).

a $\begin{bmatrix} 20 \\ 7 \\ 1 \\ 4 \end{bmatrix}$ + b $\begin{bmatrix} 17 \\ 14 \\ -8 \\ -3 \end{bmatrix}$ + c $\begin{bmatrix} -3 \\ -3 \\ 2 \\ 1 \end{bmatrix}$ + d $\begin{bmatrix} 0 \\ -1 \\ 1 \\ 1 \end{bmatrix}$

Using the vectors solved before, we can see the final result.

$\begin{bmatrix} 20a + 17b - 3c \\ a + 14b - 3c - d \\ a - 8b + 2c + d \\ 4a - 3b + c + d \end{bmatrix}$

This is our formula.

